Extracting Application and Menu Events from UI Tutorial Videos

### 2015-05-12 paper/main.tex
Revised the main file.

The proliferation of video tutorials on using different software raises new research questions on
how to improve users' ability to learn from them~\cite{matejka_ambient_2011,
pongnumkul_pause-and-play_2011}.  One strength of video tutorials is that users can directly observe
expert use. Grossman \& Fitzmaurice showed that short, 10-25 second contextually-related videos were
more effective to help users accomplish tasks than traditional text-based
tutorials~\cite{grossman_toolclips_2010}.  Unfortunately, navigation issues could lead to
misunderstanding of content, and users may not be able to keep up with the pace of instruction.
Researchers working with longer, 2-3 minute task-oriented tutorials found that users performed
better with text tutorials than videos because users could not work at the same pace as the
video~\cite{grabler_generating_2009}. In general, research on instructional videos shows a need to
segment videos to emphasize various steps. Our goal in this project is to explore how to segment
existing videos by using computer vision techniques.

By examining pixels, many have succeeded at reverse engineering user interfaces (UIs) and
interactions with purely visual information, independent of underlying application knowledge.  For
instance, Sikuli~\cite{yeh_sikuli_2009} uses computer vision to identify GUI elements from screen
captures by using template matching for small UI elements and voting based on invariant local
features for large elements. Sikuli allows users to search for documentation related to screenshots
of UI elements, and to script GUI tasks using visual cues.

Pause-and-Play extends the template matching approach of Sikuli to detect tool-selection UI events
in compressed, low-resolution video tutorials.  The system takes a tool palette template and tool
icons as input and produces a metadata file of tool changes with timecodes.  With template matching
at multiple resolutions, detection is invariant to recording resolution and some camera
effects~\cite{pongnumkul_pause-and-play_2011}.

% Can cite these but we need more space for actual content: ~\cite{dixon_content_2011}~\cite{dixon_general-purpose_2012}.
Prefab identifies GUI elements with GUI-specific visual features~\cite{dixon_prefab_2010}, which
enables overlay of advanced interactions on existing interfaces.  Prefab identifies interface
elements using two strategies: exact matches of prototype pixels, and models of the background and
pixel differences to identify foreground interface elements. Waken~\cite{banovic_waken_2012}
contributes to this research by discovering cursor shapes, clicking actions, icons, tooltips and
menus, all by examining the difference between frames of UI video.  They do this by making
assumptions about how the image of icons change during hover and press actions.  Detection of the
cursor and icons in later frames is performed by template matching.  We rely on techniques similar
to those in Waken to establish when menus have been activated and menu items selected.

@@

Recent work~\cite{banovic_waken_2012} has shown how to extract templates for cursors and icons by
monitoring pixel differences across frames.  While there has been some initial efforts
towards recognizing tooltips and menus~\cite{banovic_waken_2012}, this task still lacks a
satisfying solution, given that reliable techniques to identify menu-based events could enable
automatic assessment of usability from UI video.

Detecting menus based on pixel difference between frames poses three challenges.  First, compressed 
video contains frame-to-frame noise that clouds the signal of a change in UI appearance, 
particularly if UI widgets tend to fade in or out gradually.  Second, these menus occlude
heterogeneous backgrounds (photos in Photoshop, colorful code in the IDE), making the difference
at each individual pixel highly-varied relative to its neighbors when a new object appears in the
foreground, even if it is of homogeneous color.  Third, depending on the operating system, 
the edges dividing menus from ribbons and borders in the main UI may be subtle compared to the
internal edges in a menu caused by text or internal horizontal lines that separate groups of options.
This makes it impossible to apply traditional segmentation techniques to extract menu regions from 
static images.

@@

Our pipeline comprises three major steps.  First, we perform \emph{temporal smoothing}
of frame-to-frame pixel differences by averaging pixel differences over a window of size 10 
(Figure~\ref{fig:pipeline2}).  This reduces noise due to compression and enables us to detect
menus that fade in instead of appearing instantly.  10 frames is one-third of a second in a $29.97fps$ video, 
long enough to capture the entire fade-in of a menu in Windows 7 (3-5 frames)
and short enough to avoid capturing multiple menu actions in the same window.  Second, we \emph{coalesce
pixel differences} into a blob that spans the menu.  We perform heavy smoothing to remove
discontinuities in edge segments and bridge blobs from menu text (Figure~\ref{fig:pipeline3}).
We binarize the smoothed image at a heuristically-determined threshold (Figure~\ref{fig:pipeline4}).
Finally, we use existing methods for contour detection to trace the blob outline (Figure~\ref{fig:pipeline5}),
from which we take a rectangular bounding box as a region proposal (Figure~\ref{fig:pipeline6}).
We ignore all blobs that are smaller than 5000 pixels (around $70x70$ pixels square) to filter
out false positives from noisy regions of the screen.

@@

The quality of our region proposals is promising.  Our technique locates region proposals that
overlap with the ground truth in 30/33 (91\%) of the test events.  We see an average 
intersection-over-union of 74\%.  Our technique produced only 2 false positive bounding rectangles 
for all menu appearances.  We note that the strength of our technique varies depending
on operating system.  For example, our system produced low-overlap region proposals
on a computer like Windows 2000, where the menu and application background color were likely
indistinguishable to our difference detection methods (Figure~\ref{fig:proposal5}, Figure~\ref{fig:proposal6}).

@@

We have described a classifier that can classify video frames, and a pipeline for extracting menus
from videos. Future would would ideally integrate both steps in an overall data mining pipeline by
using class knowledge to develop more specific templates for menu extraction.

>>>>

In recent years, video tutorials for learning to use software have proliferated online.  
This has attracted interest in improving users' ability to learn from them~\cite{matejka_ambient_2011,
pongnumkul_pause-and-play_2011}.  Video tutorials appear to be beneficial for
some learning tasks.  Grossman \& Fitzmaurice showed that short, 10-25 second contextually-related videos were
more effective in helping users accomplish tasks than traditional text-based
tutorials~\cite{grossman_toolclips_2010}.

Unfortunately, navigation issues lead to misunderstandings, and users can't always keep up with the pace of instruction.
Researchers found that when provided with 2-3 minute task-oriented tutorials, users perform
better with text tutorials than videos because users cannot work at the same pace as the
video~\cite{grabler_generating_2009}. This work points to a need to better segment, index, and 
enable navigation of videos to improve users' abilities to adapt to the pace of instruction.
In our project, we investigate techniques for \emph{understanding applications and menu events in UI tutorial videos through
computer vision}, with the intent to enable new methods of segmenting, indexing and navigating tutorial videos.

We build upon past work that visually reverse engineers user interfaces (UIs).
By examining pixels, many have succeeded at reverse engineering UIs and
interactions with purely visual information, independent of underlying application knowledge.  For
instance, Sikuli~\cite{yeh_sikuli_2009} uses computer vision to identify GUI elements from screen
captures with template matching for small UI elements and voting based on invariant local
features for large elements. Prefab identifies GUI elements with GUI-specific visual features~\cite{dixon_prefab_2010}
to provide an overlay of advanced interactions on existing interfaces.  The system identifies interface
elements using two strategies: exact matches of prototype pixels, and models of the background and
pixel differences to identify foreground interface elements.

Perhaps most related, Waken~\cite{banovic_waken_2012}
contributes to this research direction by discovering cursor shapes, clicking actions, icons, tooltips and
menus, all by examining the difference between frames of UI video.  They do this by making
assumptions about how the image of icons change during hover and press actions.  Detection of the
cursor and icons in later frames is performed by template matching.  We rely on techniques similar
to those in Waken to detect when and where menus appear in UI videos.

@@

Recent work~\cite{banovic_waken_2012} has shown how to extract templates for cursors and icons by
monitoring pixel differences across frames of UI videos.  While there have been some initial efforts
towards recognizing tooltips and menus~\cite{banovic_waken_2012}, this task still lacks a
satisfying solution, given that reliable techniques to identify menu-based events could enable
automatic assessment of usability from UI video and provide new ways of indexing tutorial videos
based on complex user interface events.

It is challenging to detect menus based on pixel difference between frames for three reasons.  First, compressed 
video contains frame-to-frame noise that clouds the signal of a change in UI appearance, 
particularly if UI widgets tend to fade in or out gradually.  Second, when menus appear, they occlude
heterogeneous backgrounds (photos in Photoshop, colorful code in the IDE), making the difference
at each individual pixel highly-varied relative to its neighbors when a new object appears in the
foreground, even if it is of homogeneous color.  Third, depending on the operating system, 
the edges dividing menus from ribbons and borders in the main UI may be subtle compared to the
internal edges in a menu caused by text or internal horizontal lines that separate groups of options.
This makes it impossible to apply traditional segmentation techniques to extract menu regions from 
static images.

@@

Our pipeline comprises three major steps.  First, we perform \emph{temporal smoothing}
of frame-to-frame pixel differences by averaging pixel differences over a window of size 10 
(Figure~\ref{fig:pipeline2}).  This reduces noise due to compression and enables us to detect
menus that fade in instead of appearing instantly.  10 frames is one-third of a second in a $29.97fps$ video, 
long enough to capture the entire fade-in of a menu in Windows 7 (3-5 frames)
and short enough to avoid capturing multiple menu actions in the same window.  Second, we \emph{coalesce
pixel differences} into a blob that spans the menu.  We perform heavy smoothing to remove
discontinuities in edge segments and bridge blobs from menu text (Figure~\ref{fig:pipeline3}).
We binarize the smoothed image at a heuristically-determined threshold (Figure~\ref{fig:pipeline4}).
Finally, we use existing methods for \emph{contour detection} to trace the blob outline (Figure~\ref{fig:pipeline5}),
from which we take a rectangular bounding box as a region proposal (Figure~\ref{fig:pipeline6}).
We ignore all blobs that are smaller than 5000 pixels (around $70x70$ pixels square) to filter
out false positives from noisy regions of the screen.

@@

The quality of our region proposals is promising.  Our technique locates region proposals that
overlap with the ground truth in 30/33 (91\%) of the test events.  We see an average 
intersection-over-union of 74\%.  Our technique produced only 2 false positive bounding rectangles 
for all menu appearances.  We note that the strength of our technique varies by
operating system.  For example, our system produced low-overlap region proposals
on a Windows 2000-style interface, where the menu and application background color were likely
indistinguishable to our difference detection methods (Figure~\ref{fig:proposal5}, Figure~\ref{fig:proposal6}).

@@

We have described a classifier that can classify video frames, and a pipeline for extracting menus
from videos. Future work would ideally integrate both steps in an overall data mining pipeline by
using class knowledge to develop more specific templates for menu extraction.
